---
title: "Quick statistic analysis"
output: pdf_document
---

# Introduction

Quick statistic based and univariate audit of the different columns

This US Census dataset contains detailed but anonymised information for approximately 300,000 people.

## Metadata

This data was extracted from the census bureau database found at
http://www.census.gov/ftp/pub/DES/www/welcome.html
Donor: Terran Lane and Ronny Kohavi
      Data Mining and Visualization
      <!-- Silicon Graphics. -->
      e-mail: terran@ecn.purdue.edu, ronnyk@sgi.com for questions.

The data was split into train/test in approximately 2/3, 1/3
proportions using MineSet's MIndUtil mineset-to-mlc.

Prediction task is to determine the income level for the person
represented by the record.  Incomes have been binned at the $50K
level to present a binary classification problem, much like the
original UCI/ADULT database.  The goal field of this data, however,
was drawn from the "total person income" field rather than the
"adjusted gross income" and may, therefore, behave differently than the
orginal ADULT goal field.

More information detailing the meaning of the attributes can be
found in http://www.bls.census.gov/cps/cpsmain.htm
To make use of the data descriptions at this site, the following mappings
to the Census Bureau's internal database column names will be needed:

|Data descriptions|Database column names|
|----|-----|
|age | AAGE|
|class of worker | ACLSWKR|
|industry code | 	ADTIND|
|occupation code | ADTOCC|
|adjusted gross income | AGI|
|education | 	AHGA|
|wage per hour | 	AHRSPAY|
|enrolled in edu inst last wk | AHSCOL|
|marital status | AMARITL|
|major industry code | AMJIND|
|major occupation code | AMJOCC|
|mace | ARACE|
|hispanic Origin | AREORGN|
|sex | ASEX|
|member of a labor union | AUNMEM|
|reason for unemployment | AUNTYPE|
|full or part time employment stat | AWKSTAT|
|capital gains | 	CAPGAIN|
|capital losses | CAPLOSS|
|divdends from stocks | DIVVAL|
|federal income tax liability | FEDTAX|
|tax filer status | FILESTAT|
|region of previous residence | GRINREG|
|state of previous residence | GRINST|
|detailed household and family stat | HHDFMX|
|detailed household summary in household | HHDREL|
|instance weight | MARSUPWT|
|migration code-change in msa | MIGMTR1|
|migration code-change in reg | MIGMTR3|
|migration code-move within reg | MIGMTR4|
|live in this house 1 year ago | MIGSAME|
|migration prev res in sunbelt | MIGSUN|
|num persons worked for employer | NOEMP|
|family members under 18 | PARENT|
|total person earnings | PEARNVAL|
|country of birth father | PEFNTVTY|
|country of birth mother | PEMNTVTY|
|country of birth self | PENATVTY|
|citizenship | 	PRCITSHP|
|total person income | PTOTVAL|
|own business or self employed | SEOTR|
|taxable income amount | TAXINC|
|fill inc questionnaire for veteran's admin | VETQVA|
|veterans benefits | VETYN|
|weeks worked in year | WKSWORK|


Basic statistics for this data set:

Number of instances data = 199523

  Duplicate or conflicting instances : 46716
  
Number of instances in test = 99762

  Duplicate or conflicting instances : 20936
  
Class probabilities for income-projected.test file

Probability for the label '- 50000' : 93.80%

Probability for the label '50000+' : 6.20%

Majority accuracy: 93.80% on value - 50000

Number of attributes = 40 (continuous : 7 nominal : 33)

```{r}
# import the learning data set into da dataframe
df <- read.csv('data/census_income_learn.csv', header = FALSE)
label <- df[[42]] == " 50000+."
sum(label)/length(label)
```

We are going to look at column one by one. At first we can see that the data have a lots of nominal featuers and the class is hightly unbalanced.

# Analysis

## Age

### Metadata
91 distinct values for attribute #0 (age) continuous

### Statistics

```{r}
ages <- df[[1]]
summary(ages)
plot(density(ages))
boxplot(ages, ages[label], ages[!label], horizontal = TRUE, names=c("global", "gt50000", "lt50000"))
```

We can see a difference between the distribution of people with an imcome greater than 50K.

## Class of worker

### Metadata
9 distinct values for attribute #1 (class of worker) nominal

class of worker: Not in universe, Federal government, Local government, Never worked, Private, Self-employed-incorporated, Self-employed-not incorporated, State government, Without pay.

### Statistics

```{r}
class.of.worker <- df[[2]]
par(mar=c(5.1, 13 ,4.1 ,2.1))
barplot(sort(prop.table(summary(class.of.worker[label]))),las=1, horiz = TRUE)
barplot(sort(prop.table(summary(class.of.worker[!label]))),las=1, horiz = TRUE)
```

private, Self-employed-incorporated, Not in universe, gouvernements (concat of 3 gouv) looks to have  

## Detailed industry recode

### Metadata
52 distinct values for attribute #2 (detailed industry recode) nominal

detailed industry recode: 0, 40, 44, 2, 43, 47, 48, 1, 11, 19, 24, 25, 32, 33, 34, 35, 36, 37, 38, 39, 4, 42, 45, 5, 15, 16, 22, 29, 31, 50, 14, 17, 18, 28, 3, 30, 41, 46, 51, 12, 13, 21, 23, 26, 6, 7, 9, 49, 27, 8, 10, 20.

### Statistics

```{r, fig.height = 10}
industry.code <- as.factor(df[[3]])
par(mar=c(5.1, 5 ,4.1 ,2.1))
barplot(prop.table(summary(industry.code[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(industry.code[!label])),las=1, horiz = TRUE)
```

0 is mostly for - 50000
we can diferenciate 0 and non 0

## Detailed occupation recode (ignored in model)

### Metadata
47 distinct values for attribute #3 (detailed occupation recode) nominal

detailed occupation recode: 0, 12, 31, 44, 19, 32, 10, 23, 26, 28, 29, 42, 40, 34, 14, 36, 38, 2, 20, 25, 37, 41, 27, 24, 30, 43, 33, 16, 45, 17, 35, 22, 18, 39, 3, 15, 13, 46, 8, 21, 9, 4, 6, 5, 1, 11, 7.

### Statistics

```{r, fig.height = 10}
occupation.code <- as.factor(df[[4]])
par(mar=c(5.1, 5 ,4.1 ,2.1))
barplot(prop.table(summary(occupation.code[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(occupation.code[!label])),las=1, horiz = TRUE)
```

Same observation the big diference is the proportion of 0

## Education

### Metadata
17 distinct values for attribute #4 (education) nominal

education: Children, 7th and 8th grade, 9th grade, 10th grade, High school graduate, 11th grade, 12th grade no diploma, 5th or 6th grade, Less than 1st grade, Bachelors degree(BA AB BS), 1st 2nd 3rd or 4th grade, Some college but no degree, Masters degree(MA MS MEng MEd MSW MBA), Associates degree-occup /vocational, Associates degree-academic program, Doctorate degree(PhD EdD), Prof school degree (MD DDS DVM LLB JD).

### Statistics

```{r, fig.height = 8}
education <- as.factor(df[[5]])
par(mar=c(5.1, 19,4.1 ,2.1))
barplot(prop.table(summary(education[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(education[!label])),las=1, horiz = TRUE)
```

there is two gategory:

  * Before Hight school: Children, 7th and 8th grade, 9th grade, 10th grade, High school graduate, 11th grade, 12th grade no diploma, 5th or 6th grade, Less than 1st grade, 1st 2nd 3rd or 4th grade 

  * Post graduate: Bachelors degree(BA AB BS), Masters degree(MA MS MEng MEd MSW MBA), Associates degree-occup /vocational, Associates degree-academic program, Doctorate degree(PhD EdD), Prof school degree (MD DDS DVM LLB JD).

Some college but no degree is in the same proportion in both categories

## Wage per hour

### Metadata
1240 distinct values for attribute #5 (wage per hour) continuous

wage per hour: continuous.
### Statistics

```{r}
wage.per.hour <- df[[6]]
summary(wage.per.hour[label])
summary(wage.per.hour[!label])
par(mar=c(5.1, 5,4.1 ,2.1))
plot(density(wage.per.hour))
boxplot(wage.per.hour[label], wage.per.hour[!label], horizontal = TRUE, names=c("gt50000", "lt50000"))
```

This variable should be combine with another one. 

## Enroll in edu inst last wk (ignored in model)

### Metadata

3 distinct values for attribute #6 (enroll in edu inst last wk) nominal

enroll in edu inst last wk: Not in universe, High school, College or university.

### Statistics

```{r}
enroll.in.edu.inst.last.wk <- df[[7]]
par(mar=c(5.1, 10 ,4.1 ,2.1))
barplot(prop.table(summary(enroll.in.edu.inst.last.wk[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(enroll.in.edu.inst.last.wk[!label])),las=1, horiz = TRUE)
```

This variable doesn't very important we may ignore it at first.

## Marital stat

### Metadata
7 distinct values for attribute #7 (marital stat) nominal

marital stat: Never married, Married-civilian spouse present, Married-spouse absent, Separated, Divorced, Widowed, Married-A F spouse present.

### Statistics

```{r}
marital.stat <- df[[8]]
par(mar=c(5.1, 13 ,4.1 ,2.1))
barplot(prop.table(summary(marital.stat[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(marital.stat[!label])),las=1, horiz = TRUE)
```

Never married, Married-civillian spouse present, others

## Major industry code (ignored in model)

### Metadata
24 distinct values for attribute #8 (major industry code) nominal

major industry code: Not in universe or children, Entertainment, Social services, Agriculture, Education, Public administration, Manufacturing-durable goods, Manufacturing-nondurable goods, Wholesale trade, Retail trade, Finance insurance and real estate, Private household services, Business and repair services, Personal services except private HH, Construction, Medical except hospital, Other professional services, Transportation, Utilities and sanitary services, Mining, Communications, Hospital services, Forestry and fisheries, Armed Forces.

### Statistics

```{r, fig.height = 10}
major.industry.code <- df[[9]]
par(mar=c(5.1, 15 ,4.1 ,2.1))
barplot(prop.table(summary(major.industry.code[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(major.industry.code[!label])),las=1, horiz = TRUE)

# compare major industry code not in universe and industry code 0
major.industry.code == ' Not in universe or children' && industry.code == 0
```

This field is a duplicate from industry code

## Major occupation code

### Metadata
15 distinct values for attribute #9 (major occupation code) nominal

major occupation code: Not in universe, Professional specialty, Other service, Farming forestry and fishing, Sales, Adm support including clerical, Protective services, Handlers equip cleaners etc , Precision production craft & repair, Technicians and related support, Machine operators assmblrs & inspctrs, Transportation and material moving, Executive admin and managerial, Private household services, Armed Forces.

### Statistics

```{r}
major.occupation.code <- df[[10]]
par(mar=c(5.1, 16 ,4.1 ,2.1))
barplot(prop.table(summary(major.occupation.code[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(major.occupation.code[!label])),las=1, horiz = TRUE)

# compare major occupation code not in universe and occupation code 0
major.occupation.code == ' Not in universe' && occupation.code == 0
# compare major occupation code not in universe and weeks worked in year
major.occupation.code == " Not in universe" && df[[40]] == 0
```


we can create a dummie variable for Sales, Professional specialty, Executive admin and managerial

## Race

### Metadata
5 distinct values for attribute #10 (race) nominal

race: White, Black, Other, Amer Indian Aleut or Eskimo, Asian or Pacific Islander.
### Statistics

```{r}
race <- df[[11]]
par(mar=c(5.1,12,4.1,2.1))
barplot(prop.table(summary(race[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(race[!label])),las=1, horiz = TRUE)
```

we can divide in 2 categries:

* White, Asian.

* Black, Other, Amer Indian Aleut or Eskimo or Pacific Islander

## Hispanic origin

hispanic origin: Mexican (Mexicano), Mexican-American, Puerto Rican, Central or South American, All other, Other Spanish, Chicano, Cuban, Do not know, NA.

### Metadata
10 distinct values for attribute #11 (hispanic origin) nominal

### Statistics

```{r}
hispanic.origin <- df[[12]]
par(mar=c(5.1,12,4.1,2.1))
barplot(prop.table(summary(hispanic.origin[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(hispanic.origin[!label])),las=1, horiz = TRUE)
```

we can be merge with race in the same groupe as Black, Other, Amer Indian Aleut or Eskimo or Pacific Islander.

## Sex

### Metadata
2 distinct values for attribute #12 (sex) nominal

sex: Female, Male.

### Statistics

```{r}
sex <- df[[13]]
par(mar=c(5.1,5,4.1,2.1))
barplot(prop.table(summary(sex[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(sex[!label])),las=1, horiz = TRUE)
```

We can keep this 2 variable

## Member of a labor union (ignored in model)

### Metadata
3 distinct values for attribute #13 (member of a labor union) nominal

member of a labor union: Not in universe, No, Yes.

### Statistics

```{r}
member.labor.union <- df[[14]]
par(mar=c(5.1,8,4.1,2.1))
barplot(prop.table(summary(member.labor.union[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(member.labor.union[!label])),las=1, horiz = TRUE)
```

This variable doesn't very important we may ignore it at first.

## Reason for unemployment (ignored in model)

### Metadata
6 distinct values for attribute #14 (reason for unemployment) nominal

reason for unemployment: Not in universe, Re-entrant, Job loser - on layoff, New entrant, Job leaver, Other job loser.

### Statistics

```{r}
reason.unemployement <- df[[15]]
par(mar=c(5.1,10,4.1,2.1))
barplot(prop.table(summary(reason.unemployement[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(reason.unemployement[!label])),las=1, horiz = TRUE)
```

This variable doesn't very important we may ignore it at first.

## Full or part time employment status

### Metadata
8 distinct values for attribute #15 (full or part time employment status) nominal

full or part time employment status: Children or Armed Forces, Full-time schedules, Unemployed part- time, Not in labor force, Unemployed full-time, PT for non-econ reasons usually FT, PT for econ reasons usually PT, PT for econ reasons usually FT.

### Statistics

```{r}
full.part.employment.stat <- df[[16]]
par(mar=c(5.1,15,4.1,2.1))
barplot(prop.table(summary(full.part.employment.stat[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(full.part.employment.stat[!label])),las=1, horiz = TRUE)
```

we can divide in 2 categries:

* Children or Armed Forces, Not in labor force
* Full-time schedules

## Capital gains

### Metadata
132 distinct values for attribute #16 (capital gains) continuous

capital gains: continuous.

### Statistics

```{r}
capital.gains <- df[[17]]
summary(capital.gains[label])
summary(capital.gains[!label])
par(mar=c(5.1, 5,4.1 ,2.1))
plot(density(capital.gains[label]))
plot(density(capital.gains[!label]))
boxplot(capital.gains[label], capital.gains[!label], horizontal = TRUE, names=c("gt50000", "lt50000"))
```

We may combine with other variables

## Capital losses

### Metadata
113 distinct values for attribute #17 (capital losses) continuous

capital losses: continuous.

### Statistics

```{r}
capital.losses <- df[[18]]
summary(capital.losses[label])
summary(capital.losses[!label])
par(mar=c(5.1, 5,4.1 ,2.1))
plot(density(capital.losses[label]))
plot(density(capital.losses[!label]))
boxplot(capital.losses[label], capital.losses[!label], horizontal = TRUE, names=c("gt50000", "lt50000"))
```

We may combine with other variables

## Dividends from stocks

### Metadata
1478 distinct values for attribute #18 (dividends from stocks) continuous

dividends from stocks: continuous.

### Statistics

```{r}
dividends.stocks <- df[[19]]
summary(dividends.stocks[label])
summary(dividends.stocks[!label])
par(mar=c(5.1, 5,4.1 ,2.1))
plot(density(dividends.stocks[label]))
plot(density(dividends.stocks[!label]))
boxplot(dividends.stocks[label], dividends.stocks[!label], horizontal = TRUE, names=c("gt50000", "lt50000"))
```

We may combine with other variables

## Tax filer status

### Metadata
6 distinct values for attribute #19 (tax filer status) nominal

tax filer status: Nonfiler, Joint one under 65 & one 65+, Joint both under 65, Single, Head of household, Joint both 65+.

### Statistics

```{r}
tax.filer.stat <- df[[20]]
par(mar=c(5.1,13,4.1,2.1))
barplot(prop.table(summary(tax.filer.stat[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(tax.filer.stat[!label])),las=1, horiz = TRUE)
```

2 categories:

* Nonfiler

* Joint both under 65

## Region of previous residence (ignored in model)

### Metadata
6 distinct values for attribute #20 (region of previous residence) nominal

region of previous residence: Not in universe, South, Northeast, West, Midwest, Abroad.

### Statistics 

```{r}
region.previous.residence <- df[[21]]
par(mar=c(5.1,8,4.1,2.1))
barplot(prop.table(summary(region.previous.residence[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(region.previous.residence[!label])),las=1, horiz = TRUE)
```

This variable doesn't very important we may ignore it at first.

## State of previous residence (ignored in model)

### Metadata
51 distinct values for attribute #21 (state of previous residence) nominal

state of previous residence: Not in universe, Utah, Michigan, North Carolina, North Dakota, Virginia, Vermont, Wyoming, West Virginia, Pennsylvania, Abroad, Oregon, California, Iowa, Florida, Arkansas, Texas, South Carolina, Arizona, Indiana, Tennessee, Maine, Alaska, Ohio, Montana, Nebraska, Mississippi, District of Columbia, Minnesota, Illinois, Kentucky, Delaware, Colorado, Maryland, Wisconsin, New Hampshire, Nevada, New York, Georgia, Oklahoma, New Mexico, South Dakota, Missouri, Kansas, Connecticut, Louisiana, Alabama, Massachusetts, Idaho, New Jersey.

### Statistics

```{r, fig.height = 10}
state.previous.residence <- gsub("?",NA,df[[22]], fixed = TRUE)
par(mar=c(5.1,13,4.1,2.1))
barplot(sort(prop.table(summary.factor(state.previous.residence[label])))[!names(sort(prop.table(summary.factor(state.previous.residence[label])))) %in% c("NA's", " Not in universe")],las=1, horiz = TRUE)
barplot(sort(prop.table(summary.factor(state.previous.residence[!label])))[!names(sort(prop.table(summary.factor(state.previous.residence[!label])))) %in% c("NA's", " Not in universe")],las=1, horiz = TRUE)
#ratio of Not in the universe and not applicable
prop.table(summary.factor(state.previous.residence[label]))[names(summary.factor(state.previous.residence[label])) %in% c("NA's", " Not in universe")]
prop.table(summary.factor(state.previous.residence[!label]))[names(summary.factor(state.previous.residence[!label])) %in% c("NA's", " Not in universe")]
```

This variable doesn't very important we may ignore it at first.

## Detailed household and family stat

### Metadata
38 distinct values for attribute #22 (detailed household and family stat) nominal

detailed household and family stat: Child <18 never marr not in subfamily, Other Rel <18 never marr child of subfamily RP, Other Rel <18 never marr not in subfamily, Grandchild <18 never marr child of subfamily RP, Grandchild <18 never marr not in subfamily, Secondary individual, In group quarters, Child under 18 of RP of unrel subfamily, RP of unrelated subfamily, Spouse of householder, Householder, Other Rel <18 never married RP of subfamily, Grandchild <18 never marr RP of subfamily, Child <18 never marr RP of subfamily, Child <18 ever marr not in subfamily, Other Rel <18 ever marr RP of subfamily, Child <18 ever marr RP of subfamily, Nonfamily householder, Child <18 spouse of subfamily RP, Other Rel <18 spouse of subfamily RP, Other Rel <18 ever marr not in subfamily, Grandchild <18 ever marr not in subfamily, Child 18+ never marr Not in a subfamily, Grandchild 18+ never marr not in subfamily, Child 18+ ever marr RP of subfamily, Other Rel 18+ never marr not in subfamily, Child 18+ never marr RP of subfamily, Other Rel 18+ ever marr RP of subfamily, Other Rel 18+ never marr RP of subfamily, Other Rel 18+ spouse of subfamily RP, Other Rel 18+ ever marr not in subfamily, Child 18+ ever marr Not in a subfamily, Grandchild 18+ ever marr not in subfamily, Child 18+ spouse of subfamily RP, Spouse of RP of unrelated subfamily, Grandchild 18+ ever marr RP of subfamily, Grandchild 18+ never marr RP of subfamily, Grandchild 18+ spouse of subfamily RP.

### Statistics

```{r, fig.height = 10}
household.family.stat <- df[[23]]
par(mar=c(5.1,19,4.1,2.1))
barplot(prop.table(summary(household.family.stat[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(household.family.stat[!label])),las=1, horiz = TRUE)
```

2 categories:

* householders

* non householders

## Detailed household summary in household (ignored in model)

### Metadata
8 distinct values for attribute #23 (detailed household summary in household) nominal

detailed household summary in household: Child under 18 never married, Other relative of householder, Nonrelative of householder, Spouse of householder, Householder, Child under 18 ever married, Group Quarters- Secondary individual, Child 18 or older.

### Statistics

```{r}
household.summary.household <- df[[24]]
par(mar=c(5.1,16,4.1,2.1))
barplot(prop.table(summary(household.summary.household[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(household.summary.household[!label])),las=1, horiz = TRUE)
```

This variable doesn't very important we may ignore it at first because redundant with the previous one.

## Instance weight (ignored in model)

### Metadata

| instance weight: ignore.
instance weight: continuous.

### Statistics

```{r}
instance.weight <- df[[25]]
summary(instance.weight[label])
summary(instance.weight[!label])
plot(density(instance.weight[label]))
plot(density(instance.weight[!label]))
boxplot(instance.weight[label], instance.weight[!label], horizontal = TRUE, names=c("gt50000", "lt50000"))
```

This variable doesn't very important we may ignore it at first.

## Migration code-change in msa (ignored in model)

### Metadata

9 distinct values for attribute #25 (migration code-change in reg) nominal

migration code-change in msa: Not in universe, Nonmover, MSA to MSA, NonMSA to nonMSA, MSA to nonMSA, NonMSA to MSA, Abroad to MSA, Not identifiable, Abroad to nonMSA.

### Statistics

```{r}
migration.code.msa <- gsub("?",NA,df[[26]], fixed = TRUE)
par(mar=c(5.1,13,4.1,2.1))
barplot(prop.table(summary.factor(migration.code.msa[label])),las=1, horiz = TRUE)
barplot(prop.table(summary.factor(migration.code.msa[!label])),las=1, horiz = TRUE)
prop.table(summary.factor(migration.code.msa[label]))[names(prop.table(summary.factor(migration.code.msa[label]))) %in% c("NA's", " Nonmover")]
prop.table(summary.factor(migration.code.msa[!label]))[names(prop.table(summary.factor(migration.code.msa[!label]))) %in% c("NA's", " Nonmover")]
```

This variable doesn't very important we may ignore it at first.

## Migration code-change in reg (ignored in model)

### Metadata

9 distinct values for attribute #25 (migration code-change in reg) nominal

migration code-change in reg: Not in universe, Nonmover, Same county, Different county same state, Different state same division, Abroad, Different region, Different division same region.

### Statistics

```{r}
migration.code.change.reg <- gsub("?",NA,df[[27]], fixed = TRUE)
par(mar=c(5.1,13,4.1,2.1))
barplot(prop.table(summary.factor(migration.code.change.reg[label])),las=1, horiz = TRUE)
barplot(prop.table(summary.factor(migration.code.change.reg[!label])),las=1, horiz = TRUE)
#ratio of Nonmover and not applicable
prop.table(summary.factor(migration.code.change.reg[label]))[names(prop.table(summary.factor(migration.code.change.reg[label]))) %in% c("NA's", " Nonmover")]
prop.table(summary.factor(migration.code.change.reg[!label]))[names(prop.table(summary.factor(migration.code.change.reg[!label]))) %in% c("NA's", " Nonmover")]
```

This variable doesn't very important we may ignore it at first.

## Migration code-move within reg (ignored in model)

### Metadata

10 distinct values for attribute #26 (migration code-move within reg) nominal

migration code-move within reg: Not in universe, Nonmover, Same county, Different county same state, Different state in West, Abroad, Different state in Midwest, Different state in South, Different state in Northeast.

### Statistics

```{r}
migration.code.move.reg <- gsub("?",NA,df[[28]], fixed = TRUE)
par(mar=c(5.1,13,4.1,2.1))
barplot(prop.table(summary.factor(migration.code.move.reg[label])),las=1, horiz = TRUE)
barplot(prop.table(summary.factor(migration.code.move.reg[!label])),las=1, horiz = TRUE)
#ratio of Nonmover and not applicable
prop.table(summary.factor(migration.code.move.reg[label]))[names(prop.table(summary.factor(migration.code.move.reg[label]))) %in% c("NA's", " Nonmover")]
prop.table(summary.factor(migration.code.move.reg[!label]))[names(prop.table(summary.factor(migration.code.move.reg[!label]))) %in% c("NA's", " Nonmover")]
```

This variable doesn't very important we may ignore it at first.

## Live in this house 1 year ago (ignored in model)

### Metadata

3 distinct values for attribute #27 (live in this house 1 year ago) nominal

live in this house 1 year ago: Not in universe under 1 year old, Yes, No.

### Statistics

```{r}
live.in.this.house.1y.ago <- df[[29]]
par(mar=c(5.1,13,4.1,2.1))
barplot(prop.table(summary(live.in.this.house.1y.ago[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(live.in.this.house.1y.ago[!label])),las=1, horiz = TRUE)
```

This variable doesn't very important we may ignore it at first.

## Migration prev res in sunbelt (ignored in model)

### Metadata

4 distinct values for attribute #28 (migration prev res in sunbelt) nominal

migration prev res in sunbelt: Not in universe, Yes, No.

### Statistics

```{r}
migration.in.sunbelt <- gsub("?",NA,df[[30]], fixed = TRUE)
par(mar=c(5.1,13,4.1,2.1))
barplot(prop.table(summary.factor(migration.in.sunbelt[label]))[!names(prop.table(summary.factor(migration.in.sunbelt[label]))) %in% c("NA's")],las=1, horiz = TRUE)
barplot(prop.table(summary.factor(migration.in.sunbelt[!label]))[!names(prop.table(summary.factor(migration.in.sunbelt[!label]))) %in% c("NA's")],las=1, horiz = TRUE)
#ratio of not applicable
prop.table(summary.factor(migration.in.sunbelt[label]))[names(prop.table(summary.factor(migration.in.sunbelt[label]))) %in% c("NA's")]
prop.table(summary.factor(migration.in.sunbelt[!label]))[names(prop.table(summary.factor(migration.in.sunbelt[!label]))) %in% c("NA's")]
```

This variable doesn't very important we may ignore it at first.

## Num persons worked for employer

### Metadata

7 distinct values for attribute #29 (num persons worked for employer) continuous

num persons worked for employer: continuous.

### Statistics

```{r}
num.persons.worked.employer <- df[[31]]
summary(num.persons.worked.employer[label])
summary(num.persons.worked.employer[!label])
plot(density(num.persons.worked.employer[label]))
plot(density(num.persons.worked.employer[!label]))
boxplot(num.persons.worked.employer[label], num.persons.worked.employer[!label], horizontal = TRUE, names=c("gt50000", "lt50000"))
```

We can clearly see a difference between the distribution of people with an income greater than 50K.

## Family members under 18 (ignored in model)

### Metadata

5 distinct values for attribute #30 (family members under 18) nominal

family members under 18: Both parents present, Neither parent present, Mother only present, Father only present, Not in universe.

### Statistics

```{r}
family.members.under.18 <- as.factor(df[[32]])
par(mar=c(5.1,10,4.1,2.1))
barplot(prop.table(summary(family.members.under.18[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(family.members.under.18[!label])),las=1, horiz = TRUE)
```

This variable doesn't very important we may ignore it at first, it is redundant with age and sone other variable because people with an income grater than 50 are mostly above 18.

## Country of birth father (ignored in model)

### Metadata

43 distinct values for attribute #31 (country of birth father) nominal

country of birth father: Mexico, United-States, Puerto-Rico, Dominican-Republic, Jamaica, Cuba, Portugal, Nicaragua, Peru, Ecuador, Guatemala, Philippines, Canada, Columbia, El-Salvador, Japan, England, Trinadad&Tobago, Honduras, Germany, Taiwan, Outlying-U S (Guam USVI etc), India, Vietnam, China, Hong Kong, Cambodia, France, Laos, Haiti, South Korea, Iran, Greece, Italy, Poland, Thailand, Yugoslavia, Holand-Netherlands, Ireland, Scotland, Hungary, Panama.

### Statistics

```{r, fig.height = 10}
country.birth.father <- gsub("?",NA,df[[33]], fixed = TRUE)
par(mar=c(5.1,13,4.1,2.1))
barplot(prop.table(summary.factor(country.birth.father[label]))[!names(prop.table(summary.factor(country.birth.father[label]))) %in% c("NA's")],las=1, horiz = TRUE)
barplot(prop.table(summary.factor(country.birth.father[!label]))[!names(prop.table(summary.factor(country.birth.father[!label]))) %in% c("NA's")],las=1, horiz = TRUE)
#ratio of not applicable
prop.table(summary.factor(country.birth.father[label]))[names(prop.table(summary.factor(country.birth.father[label]))) %in% c("NA's")]
prop.table(summary.factor(country.birth.father[!label]))[names(prop.table(summary.factor(country.birth.father[!label]))) %in% c("NA's")]
```

This variable doesn't very important we may ignore it at first.

## Country of birth mother (ignored in model)

### Metadata

43 distinct values for attribute #32 (country of birth mother) nominal

country of birth mother: India, Mexico, United-States, Puerto-Rico, Dominican-Republic, England, Honduras, Peru, Guatemala, Columbia, El-Salvador, Philippines, France, Ecuador, Nicaragua, Cuba, Outlying-U S (Guam USVI etc), Jamaica, South Korea, China, Germany, Yugoslavia, Canada, Vietnam, Japan, Cambodia, Ireland, Laos, Haiti, Portugal, Taiwan, Holand-Netherlands, Greece, Italy, Poland, Thailand, Trinadad&Tobago, Hungary, Panama, Hong Kong, Scotland, Iran.

### Statistics

```{r, fig.height = 10}
country.birth.mother <- gsub("?",NA,df[[34]], fixed = TRUE)
par(mar=c(5.1,13,4.1,2.1))
barplot(prop.table(summary.factor(country.birth.mother[label]))[!names(prop.table(summary.factor(country.birth.mother[label]))) %in% c("NA's")],las=1, horiz = TRUE)
barplot(prop.table(summary.factor(country.birth.mother[!label]))[!names(prop.table(summary.factor(country.birth.mother[!label]))) %in% c("NA's")],las=1, horiz = TRUE)
#ratio of not applicable
prop.table(summary.factor(country.birth.mother[label]))[names(prop.table(summary.factor(country.birth.mother[label]))) %in% c("NA's")]
prop.table(summary.factor(country.birth.mother[!label]))[names(prop.table(summary.factor(country.birth.mother[!label]))) %in% c("NA's")]
```

This variable doesn't very important we may ignore it at first.

## Country of birth self (ignored in model)

### Metadata

43 distinct values for attribute #33 (country of birth self) nominal

country of birth self: United-States, Mexico, Puerto-Rico, Peru, Canada, South Korea, India, Japan, Haiti, El-Salvador, Dominican-Republic, Portugal, Columbia, England, Thailand, Cuba, Laos, Panama, China, Germany, Vietnam, Italy, Honduras, Outlying-U S (Guam USVI etc), Hungary, Philippines, Poland, Ecuador, Iran, Guatemala, Holand-Netherlands, Taiwan, Nicaragua, France, Jamaica, Scotland, Yugoslavia, Hong Kong, Trinadad&Tobago, Greece, Cambodia, Ireland.

### Statistics

```{r, fig.height = 10}
country.birth.self <- gsub("?",NA,df[[34]], fixed = TRUE)
par(mar=c(5.1,13,4.1,2.1))
barplot(prop.table(summary.factor(country.birth.self[label]))[!names(prop.table(summary.factor(country.birth.self[label]))) %in% c("NA's")],las=1, horiz = TRUE)
barplot(prop.table(summary.factor(country.birth.self[!label]))[!names(prop.table(summary.factor(country.birth.self[!label]))) %in% c("NA's")],las=1, horiz = TRUE)
#ratio of not applicable
prop.table(summary.factor(country.birth.self[label]))[names(prop.table(summary.factor(country.birth.self[label]))) %in% c("NA's")]
prop.table(summary.factor(country.birth.self[!label]))[names(prop.table(summary.factor(country.birth.self[!label]))) %in% c("NA's")]
```

This variable doesn't very important we may ignore it at first. the only factor we can find is redundant with the race hispanic.

## Citizenship (ignored in model)

### Metadata

5 distinct values for attribute #34 (citizenship) nominal

citizenship: Native- Born in the United States, Foreign born- Not a citizen of U S , Native- Born in Puerto Rico or U S Outlying, Native- Born abroad of American Parent(s), Foreign born- U S citizen by naturalization.

### Statistics

```{r}
citizenship <- do.call(rbind ,strsplit(as.character(df[[36]]), '-'))
native.citizen <- as.factor(citizenship[,1])
status.citizen <- as.factor(citizenship[,2])
par(mar=c(5.1,18,4.1,2.1))
barplot(prop.table(summary(native.citizen[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(native.citizen[!label])),las=1, horiz = TRUE)
barplot(prop.table(summary(status.citizen[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(status.citizen[!label])),las=1, horiz = TRUE)
```

This variable doesn't very important we may ignore it at first.

## Own business or self employed (ignored in model)

### Metadata

3 distinct values for attribute #35 (own business or self employed) nominal

own business or self employed: 0, 2, 1.

### Statistics

```{r}
own.business.or.self.employed <- as.factor(df[[37]])
par(mar=c(5.1,6,4.1,2.1))
barplot(prop.table(summary(own.business.or.self.employed[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(own.business.or.self.employed[!label])),las=1, horiz = TRUE)
```

This variable doesn't very important we may ignore it at first.

## Fill inc questionnaire for veteran's admin

### Metadata

3 distinct values for attribute #36 (fill inc questionnaire for veteran's admin) nominal

fill inc questionnaire for veteran's admin: Not in universe, Yes, No.

### Statistics

```{r}
fill.inc.questionnaire.for.veterant.admin <- as.factor(df[[38]])
par(mar=c(5.1,8,4.1,2.1))
barplot(prop.table(summary(fill.inc.questionnaire.for.veterant.admin[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(fill.inc.questionnaire.for.veterant.admin[!label])),las=1, horiz = TRUE)
#ratio of not applicable
prop.table(summary(fill.inc.questionnaire.for.veterant.admin[label]))
prop.table(summary(fill.inc.questionnaire.for.veterant.admin[!label]))
```

Boolean yes/no

## Veterans benefits

### Metadata

3 distinct values for attribute #37 (veterans benefits) nominal

veterans benefits: 0, 2, 1.

### Statistics

```{r}
veterants.benefits <- as.factor(df[[39]])
par(mar=c(5.1,4,4.1,2.1))
barplot(prop.table(summary(veterants.benefits[label])),las=1, horiz = TRUE)
barplot(prop.table(summary(veterants.benefits[!label])),las=1, horiz = TRUE)
#ratio of not applicable
prop.table(summary(veterants.benefits[label]))
prop.table(summary(veterants.benefits[!label]))
```

Boolean 0 / 1 or 2

## Weeks worked in year

### Metadata

53 distinct values for attribute #38 (weeks worked in year) continuous

weeks worked in year: continuous.

### Statistics

```{r}
weeks.worked.in.year <- df[[40]]
summary(weeks.worked.in.year[label])
summary(weeks.worked.in.year[!label])
plot(density(weeks.worked.in.year[label]))
plot(density(weeks.worked.in.year[!label]))
boxplot(weeks.worked.in.year[label], weeks.worked.in.year[!label], horizontal = TRUE, names=c("gt50000", "lt50000"))
```

We can clearly see a difference between the distribution of people with an income greater than 50K.

## Year (ignored in model)

### Metadata

2 distinct values for attribute #39 (year) nominal
year: 94, 95.

### Statistics

```{r}
year <- as.factor(df[[41]])
sort(prop.table(summary(year[label])))
sort(prop.table(summary(year[!label])))
```

This variable doesn't very important we may ignore it at first.


# Summary

The column selected for the model are:

Age

Class of worker
  
* worker private: " Private"
* worker public: [" Local government", " State government", " Federal government"]

Industrie code (0, !0)

Education
  
  * no university: [" Children", " 1st 2nd 3rd or 4th grade", " Less than 1st grade", " 5th or 6th grade", " 7th and 8th grade", " 9th grade", " 10th grade", " 11th grade", " 12th grade no diploma", " High school graduate", ]
  * university degree: [" Bachelors degree(BA AB BS)", " Masters degree(MA MS MEng MEd MSW MBA)", " Associates degree-occup /vocational", " Associates degree-academic program", " Doctorate degree(PhD EdD)", " Prof school degree (MD DDS DVM LLB JD)"]
  
Wage per hour

Marital state
  
* never married: " Never married"
* Married civilian spouse present: " Married-civilian spouse present"
  
Major occupation code

* well paid occupation: [" Sales", " Professional specialty", " Executive admin and managerial"]
  
Race boolean

* white or asian: [" White", " Asian] / [" Black", " Pacific Islander", "hispanic", " Amer Indian Aleut or Eskimo"]
  
Sex as a dummie variable " Male" / " Female"

Full or part time employment stat

* not in labor force or armed forces: [" Children or Armed Forces", " Not in labor force"]
* Full time schedules: " Full-time schedules"

Capital gains

Capital losses

Dividends from stocks

Tax filer stat

* Tax nonfiler: " Nonfiler"
* Tax joint both under 65: " Joint both under 65"

Detailed household and family stat as dummie variable " householders" or not

Num persons worked for employer

Fill inc questionnaire for veteran’s admin 
  
  * [" Yes", " No"] / " Not in universe"
  
Veterans benefits

* O / [1, 2]

Weeks worked in year
